Constructive Preference Elicitation by Setwise Max-margin Learning

### 2016-01-26 paper/ijcai16.tex
edits in the abstract a# Please enter the commit message for your changes. Lines starting

\title{Constructive Preference Elicitation by Setwise Max-margin Learning}
\author{ST and AP and PV}

@@

\begin{abstract}
WRITEME
\end{abstract}

@@

@@

- importance of recommendation systems and preference elicitation

- existing approaches and their limitations

% constructive preference elicitation
In this paper we take a {\em constructive} view on preference
elicitation, enlarging its scope from the selection of items among a
set of candidates to the synthesis of entirely novel
instances. Instances are represented as combinations of basic elements
(e.g. the components of a laptop) subject to a set of constraints
(e.g. the laptop model determines the set of available CPUs). A
utility function is learned over the feature representation of an
instance, as customary in many preference elicitation approaches. The
recommendation is then made by solving a constrained optimization
problem in the space of feasible instances, guided by the learned
utility. \andrea{other constructive preference elicitation approaches?}

% setwise max-margin formulation to deal with uncertainty in user utility
Learning a user utility function requires setting a preference
elicitation strategy and dealing with uncertain and possibly
inconsistent user feedback. Bayesian preference elicitation approaches
deal with both problems by building a probability distribution on
candidate functions and asking queries maximizing informativeness
measures such as expected value of information (EVOI)~\cite{}. These
approaches are however computationally expensive and cannot scale to
fully constructive scenarios, as shown in our experimental results.
We take a space decomposition perspective and jointly learn a set of
weight vectors, each representing a candidate utility function,
maximizing diversity between the vectors and consistency with the
available feedback. These two conflicting objectives tend to generate
equally plausible alternative hypotheses for the unknown
utility. Preference elicitation works by combining weight vector
learning with instance generation, so that each iteration of the
algorithm produces two outcomes: a set of weight vectors and a set of
instances, each maximizing its score according to one of the weight
vectors.

- experimental results

- paper structure

\section{Preference Elicitation}

@@

In adherence to previous work~\cite{guo2010real,viappiani2010optimal}, we model users by means
of linear utility functions: the user's preferences are represented by a weight
vector $\vw\in\bbR^m$, and the utility of a configuration $\vx$ is given by
$\langle \vw, \vx \rangle = \sum_{z=1}^m w_z x_z$. In the remainder of the paper we
require all weights to be {\em non-negative} and {\em bounded}: the per-attribute
weights $w_z$ must lie in a (constant but otherwise arbitrary) interval
$[w^\bot_z, w^\top_z]$, with $w^\bot_z \ge 0$. Both requirements are quite
natural, and enable the translation of our core optimization problem into a
mixed-integer linear problem (as done in Section~\ref{sec:formulation}).

During learning, the actual weight vector $\vw$ is {\em unknown} to the
learning system, and must be estimated from pairwise choices elicited from the
user.\andrea{@PAOLO: qualche parola e reference per giustificare la scelta di pairwise queries?}  We consider three possible outcomes for a pairwise comparison between
two configurations $\vx$ and $\vx'$: either $\vx$ is preferred to $\vx'$ (written
$\vx \succ \vx'$), $\vx'$ is preferred to $\vx$ ($\vx \prec \vx'$), or there is
no clear preference between the two items ($\vx \approx \vx'$). We write
$\calD$ to denote the full set of answers elicited from the user. Queries are
typically selected so as to maximize some measure of informativeness of the
resulting responses.  Details on our actual query selection strategy will be
provided in Section~\ref{sec:formulation}.

@@

The goal of our setwise max-margin approach is twofold. First, for any
given set size $k\geq 1$, we want to find a {\em set} of $k$ weight
vectors $\vw^{1}, \ldots, \vw^{k}$, chosen so that all user-provided
preferences are satisfied by the largest possible margin (i.e. all
weight vectors are consistent with respect to the user responses
$\calD$, modulo inconsistencies) and so that they are maximally
diverse.  Second, we want to construct a {\em set} of $k$
configurations $\vx^{1}, \ldots, \vx^{k}$, so that each configuration
$\vx^{i}$ is the ``best'' possible option when evaluated according to
the corresponding $\vw^{i}$ and configurations are maximally diverse
among each other. These options will be later employed to formulate
the user queries.

@@

\paragraph{MILP Formulation.} This initial formulation is problematic,
as Eq.~(\ref{eq:wxconstr}) involves quadratic terms over mixed
continuous integer variables. Here we show how to reformulate the
problem as a mixed integer linear program (MILP). Our goal is to
replace Eq.~(\ref{eq:wxconstr}) with a set of linear constraints through
a suitable variable transformation.

@@

We adopt the {\em indifference-augmented} Bradley-Torr user response model
introduced in~\cite{guo2010real}. When queried about a pair of configurations,
the user replies with the true utility of the two items. According to the
classical Bradley-Torr model, which does not attempt to model {\em indifference},
the proabability that a user $\vw$ will prefer configuration $\vx^i$ over
the alternative $\vx^j$, barring indifference, amounts to:
%
$$ (1 + \exp(-\alpha \langle\vw,\vx^i - \vx^j\rangle))^{-1} $$
%
In other words, the probability grows with the difference between the utilities
of the two items. Support for indifference has been introduced
in~\cite{guo2010real} by postulating that the probability of the user being
indifferent about the two configurations $\vx^i, \vx^j$ depends on how close
the two utilities are, as in:
%
$$ \exp(\beta |\langle\vw,\vx^i - \vx^j\rangle|) $$
%
The parameters $\alpha$ and $\beta$ were set as in~\cite{guo2010real}. Please
see the latter for more details
\stefano{The above does not satisfy me at all.}

@@

\paragraph{Cognitive effort.} Contrarily to most other preference elicitation
methods, which generate a single candidate option per iteration, {\sc
SetMargin} constructs a whole {\em set} of options. We thus can not leverage
standard query selection strategies \stefano{which ones?}.

@@

\section*{Acknowledgments}

>>>>

\title{Constructive Preference Elicitation by Setwise Max-margin Learning} %with?
\author{ID xxxx}

@@

\begin{abstract}
In this paper we propose an approach to preference elicitation that is suitable to large
configuration problem domains where stated preferences can be noisy.
Our setwise max-margin method can be viewed as a generalization of max-margin learning to sets,
and can produce a set of ``diverse'' items that can be used to ask informative queries to the user.
Moreover, the approach can ensuring sparsity in the parameter space, in order to favor the assessment of utility towards combinations of weights that concentrate on just few features.
We present an optimization formulation and show, with simulations, how our approach is able to efficiently elicit the user's utility function even in very large datasets.

\end{abstract}

@@

%The problem is that of learning or assessing the preferences of the user (or decision maker).
Preferences \cite{Peintner2008} play an important role in artificial intelligence,
including fields spanning from recommender systems to automatic planning, 
from non monotonic reasoning to computational social choice and algorithmic decision theory.
The elicitation of learning of preferences is therefore a crucial task, .
The goal in preference elicitation is that of gaining enough information about the user's preferences (typically represented by an utility function $u$, unknown to the system) in order to recommend a decision.

While preference elicitation has been often cast as that of adapting using ad-hoc rules \cite{},
recently a number of researchers have proposed principled methods that elicit utility in an adaptive way \cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,braziunas-mmr:uai07,guo2010real,viappiani2010optimal}; the aims is that of focusing on learning the ``important'' part of the utility, allowing to recommend good (or even optimal) decision with only a partial information about the utility function.
One option is to be Bayesian and to maintain a probability distribution over the possible realizations of the utility's parameters \cite{chajewska2000,boutilier2002,guo2010real,viappiani2010optimal}.
Alternatively, when distribution information about the parameters is not available, one can reason about all feasible utility functions consistent with the currently known information about the user; from the user's response we are able to infer constraints on the parameters implicated in the utility model \cite{Wang2003,boutilier2006,braziunasmmr:uai07,viappiani2009}.
This latter approach has also the advantage of avoiding the computationally demanding task of maintaining distribution information and performing Bayesian updates.

% constructive preference elicitation
In this paper we take a {\em constructive} view on preference
elicitation, enlarging its scope from the selection of items among a
set of candidates to the synthesis of entirely novel
instances. Instances are represented as combinations of basic elements
(e.g. the components of a laptop) subject to a set of constraints
(e.g. the laptop model determines the set of available CPUs). A
utility function is learned over the feature representation of an
instance, as customary in many preference elicitation approaches. The
recommendation is then made by solving a constrained optimization
problem in the space of feasible instances, guided by the learned
utility.

%\andrea{other constructive preference elicitation approaches?}

% setwise max-margin formulation to deal with uncertainty in user utility
Learning a user utility function requires setting a preference
elicitation strategy and dealing with uncertain and possibly
inconsistent user feedback. Bayesian preference elicitation approaches
deal with both problems by building a probability distribution on
candidate functions and asking queries maximizing informativeness
measures such as expected value of information (EVOI)~\cite{}. These
approaches are however computationally expensive and cannot scale to
fully constructive scenarios, as shown in our experimental results.
We take a space decomposition perspective and jointly learn a set of
weight vectors, each representing a candidate utility function,
maximizing diversity between the vectors and consistency with the
available feedback. These two conflicting objectives tend to generate
equally plausible alternative hypotheses for the unknown
utility. 
Our approach to preference elicitation works by combining weight vector
learning with instance generation, so that each iteration of the
algorithm produces two outcomes: a set of weight vectors and a set of
instances, each maximizing its score according to one of the weight
vectors.

We evaluate the effectiveness of our approach by testing our elicitation method
in both synthetic and configuration problems, and comparing it to state-of-the-art
methods.

The paper is structured as follows: after discussing related works in next Section,
we present some brackground material in Section\ref{sec:background} 
and then proceed to our setwise maxmargin method in Section\ref{sec:formulationm}; 
we discuss experimental results in Section\ref{sec:experiments}
and conclude with final remarks (Section\ref{sec:conclusions}).

Preference elicitation for the customization of user interfaces \cite{gajos2005}.

@@

\section{Background}
\ref{sec:background}
%\section{Preference Elicitation}

@@

Consistently with previous work~\cite{guo2010real,viappiani2010optimal}, we model users by means
of linear utility functions: the user's preferences are represented by a weight
vector $\vw\in\bbR^m$, and the utility of a configuration $\vx$ is given by
$\langle \vw, \vx \rangle = \sum_{z=1}^m w_z x_z$. In the remainder of the paper we
require all weights to be {\em non-negative} and {\em bounded}: the per-attribute
weights $w_z$ must lie in a (constant but otherwise arbitrary) interval
$[w^\bot_z, w^\top_z]$, with $w^\bot_z \ge 0$. Both requirements are quite
natural, and enable the translation of our core optimization problem into a
mixed-integer linear problem (as done in Section~\ref{sec:formulation}).

During learning, the actual weight vector $\vw$ is {\em unknown} to the
learning system, and must be estimated from pairwise choices elicited from the
user.

%\andrea{@PAOLO: qualche parola e reference per giustificare la scelta di pairwise queries?}  
Comparison queries are especially natural, asking a user if she prefers one option x to another y;
these comparisons can be extended to choice sets of more than two options \cite{viappiani2009,viappiani2010optimal}.
Comparison and choice queries are common in conjoint analysis \cite{louviere2000,toubia2004}.

We consider three possible outcomes for a pairwise comparison between
two configurations $\vx$ and $\vx'$: either $\vx$ is preferred to $\vx'$ (written
$\vx \succ \vx'$), $\vx'$ is preferred to $\vx$ ($\vx \prec \vx'$), or there is
no clear preference between the two items ($\vx \approx \vx'$). We write
$\calD$ to denote the set of preferences (answers to comparison queries) 
elicited from the user. Queries are typically selected so as to maximize 
some measure of informativeness of the
resulting responses.  Details on our actual query selection strategy will be
provided in Section~\ref{sec:formulation}.

@@

The goal of our setwise max-margin approach is twofold. First, for any
given set size $k\geq 1$, we want to find a {\em set} of $k$ weight
vectors $\vw^{1}, \ldots, \vw^{k}$, chosen so that all user-provided
preferences are satisfied by the largest possible margin (i.e. all
weight vectors are consistent with respect to the user responses
$\calD$, modulo inconsistencies) and so that they are maximally
diverse.  Second, we want to construct a {\em set} of $k$
configurations $\vx^{1}, \ldots, \vx^{k}$, so that each configuration
$\vx^{i}$ is the ``best'' possref:keeney-mautible option when evaluated according to
the corresponding $\vw^{i}$ and configurations are maximally diverse
among each other. These options will be later employed to formulate
the user queries.

@@

\paragraph{MILP Formulation.} This initial formulation is problematic to solve,
as Eq.~(\ref{eq:wxconstr}) involves quadratic terms over mixed continuous integer variables. 
Here we show how to reformulate the
problem as a mixed integer linear program (MILP). Our goal is to
replace Eq.~(\ref{eq:wxconstr}) with a set of linear constraints through
a suitable variable transformation.

@@

We adopt the {\em indifference-augmented} Bradley-Torr user response model
introduced in~\cite{guo2010real}. When queried about a pair of configurations,
the user replies with the true utility of the two items. According to the
classical Bradley-Torr model, which does not attempt to model {\em indifference},
the probability that a user $\vw$ will prefer configuration $\vx^i$ over
the alternative $\vx^j$, barring indifference, amounts to:
%
$$ (1 + \exp(-\alpha \langle\vw,\vx^i - \vx^j\rangle))^{-1} $$
%
In other words, the probability grows with the difference between the utilities
of the two items. Support for indifference has been introduced
in~\cite{guo2010real} by postulating that the probability of the user being
indifferent about the two configurations $\vx^i, \vx^j$ depends on how close
the two utilities are, as in:
%
$$ \exp(\beta |\langle\vw,\vx^i - \vx^j\rangle|) $$
%
The parameters $\alpha$ and $\beta$ were set as in~\cite{guo2010real}. Please
see the latter for more details
\stefano{The above does not satisfy me at all.}

@@

\paragraph{Cognitive effort.} Contrarily to most other preference elicitation
methods, which generate a single candidate option per iteration, {\sc
SetMargin} constructs a whole {\em set} of options. We thus can not leverage
standard query selection strategies \stefano{which ones?}. 
\paolo{I would remove the last sentence}

@@

%\section*{Acknowledgments}
%WRITEME
